Reproducibe Research Progamming Assignment 1
========================================================

### Read raw data.
```{r}
A <- read.csv('activity.csv', as.is=T)
head(A)
tail(A)
fieldnames <- names(A)
print("field names...", quote=F)
print(fieldnames)
```

### Seconds per day, days, expected number of samples, sample count check.
```{r}
spd <- 24 * 60 / 5
days = 31 + 30
sbSamples <- spd * days
ns <- nrow(A)
print(sprintf("got (%d) == expected (%d) samples: %s", ns, sbSamples, ns == sbSamples), quote=F)
```

### Sorted?
```{r}
z <- A[order(A$date, A$interval), ]
print("data are sorted by date, interval...", quote=F)
ns == sum(z$date == A$date)
ns == sum(z$interval == A$interval)
rm(z)
```

### Field types.
```{r}
class(A$steps)
class(A$date)
class(A$interval)
```

### NA, Zero, NonZero counts (by interval).
```{r}
nAna <- sum(is.na(A$steps))
nAz <- sum(A$steps == 0, na.rm=T)
nAnz <- sum(A$steps != 0, na.rm=T)
print(sprintf("A$steps NA: %d, zero: %d, nonzero: %d", nAna, nAz, nAnz), quote=F)
```

### Save the raw steps value in another field.
```{r}
A$rawSteps <- A$steps
```

### Add a field indicating that steps is NA.
```{r}
A$naSteps <- is.na(A$steps)

## Add an empty field for imputed steps.
```{r}
A$impSteps <- NA
```

### Add a day-of-week field.
```{r}
A$DOW <- weekdays(strptime(A$date, format='%Y-%m-%d'), abbreviate=T)
A$facDOW <- factor(A$DOW, levels=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"), ordered=T)
```

### Add a weekday/weekend factor.
```{r}
A$wdwe <- factor(A$facDOW >= "Sat" & A$facDOW <= "Sun", labels=c("weekday", "weekend"))
sum(A$wdwe == "weekday")
sum(A$wdwe == "weekend")
```

### Split steps by date, interval
```{r}
splitStepsByDate <- split(A$steps, as.factor(A$date))
splitStepsByInterval <- split(A$steps, as.factor(A$interval))
length(splitStepsByDate)
length(splitStepsByInterval)
```

### Aggregate by date, interval.
```{r}
StepsByDate <- sapply(splitStepsByDate, sum, na.rm=T)
meanStepsByDate <- sapply(splitStepsByDate, mean, na.rm=T)
medianStepsByDate <- sapply(splitStepsByDate, median, na.rm=T)
StepsByInterval <- sapply(splitStepsByInterval, sum, na.rm=T)
meanStepsByInterval <- sapply(splitStepsByInterval, mean, na.rm=T)
medianStepsByInterval <- sapply(splitStepsByInterval, median, na.rm=T)
```

### Histogram of steps per day.
```{r}
hist(StepsByDate)

### Report mean, median by day
```{r}
print("Date, Mean steps, Median steps...", quote=F)
for (i in 1 : length(StepsByDate)) {
  print(sprintf("%s  mean %4.1f  median %4.1f", names(StepsByDate[i]), meanStepsByDate[i], medianStepsByDate[i]), quote=F)
}
```

### Plot steps by interval.
```{r}
plot(meanStepsByInterval, type="l", ylab="Average Steps", main="Average Steps Per 5-minute Interval")
```

### Which interval has max steps?
```{r}
maxI = 1
maxMean = meanStepsByInterval[[1]]
for (i in 2 : length(meanStepsByInterval)) {
  if (meanStepsByInterval[[i]] > maxMean) {
    maxI <- i
    maxMean <- meanStepsByInterval[[i]]
  }
}
print(sprintf("Interval %d has maximum average steps (%.1f)", maxI, maxMean), quote=F)
```

### Impute missing steps from 5-minute medians.
```{r}
nI = 0
nZimpute <- 0
nNZimpute <- 0
for (i in 1 : ns) {
  if (A$naSteps[i]) {
    nI <- nI + 1
    for (j in 1 : length(medianStepsByInterval)) {
      k <- as.integer(names(medianStepsByInterval[j]))
      m <- medianStepsByInterval[j]
        if (A$interval[i] == k) {
          if (m == 0) {
            nZimpute <- nZimpute + 1
          } else {
            nNZimpute <- nNZimpute + 1
          }
        A$impSteps <- m
        break
      }
    }
    A$impSteps[i] <- A$impSteps[i]
  } else {
    A$impSteps[i] <- A$steps[i]
  }
}
nI
nZimpute
nNZimpute
nI == sum(is.na(A$steps))
```

### Copy imputed steps into steps.
```{r}
A$steps <- A$impSteps
ns == sum(is.na(A$steps))
```

### Redo aggregation by date, interval.
```{r}
StepsByDate <- sapply(splitStepsByDate, sum, na.rm=T)
meanStepsByDate <- sapply(splitStepsByDate, mean, na.rm=T)
medianStepsByDate <- sapply(splitStepsByDate, median, na.rm=T)
StepsByInterval <- sapply(splitStepsByInterval, sum, na.rm=T)
meanStepsByInterval <- sapply(splitStepsByInterval, mean, na.rm=T)
medianStepsByInterval <- sapply(splitStepsByInterval, median, na.rm=T)
```

### Redo histogram of steps per day.
```{r}
hist(StepsByDate)
```

### Redo report mean, median by day
```{r}
print("Date, Mean steps, Median steps...", quote=F)
for (i in 1 : length(StepsByDate)) {
  print(sprintf("%s  mean %4.1f  median %4.1f", names(StepsByDate[i]), meanStepsByDate[i], medianStepsByDate[i]), quote=F)
}
```

## End.

